---
title: "Chapter2"
author: "Henry Hansen"
date: "11/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
```

## 2.1 Choices
The multiplication principle - is the number of ways of making a sequence of independent choices is the product of the number of choices at each step.The key is the "independent" aspect which means that how you make the first choice doesn't affect teh number of choices you have for the second, and so on. 

### Exercise 2.1.1
```{r}
nailPolish <- 7
fingers <- 10
nailPolish * fingers #just multiply
```

## 2.2 Counting Words
Let's try rebuilding the table from this chapter to better understand choices and how choices can change the outcome we are interested in. 

```{r}
n <- c(1:9) #create sequence of "words". Here we want to see up to 9 letter words

f <- function(n) {
    now <- 26 ^ n # calculate number of words
    wor <- factorial(26) / (factorial(26 - n)) # calculate without repeats using permutations
    perc <- round(wor/now*100,0) # calculate percentage
    matrix(c(n, now,wor,perc))# output data
    
}
out <- sapply(n, f) #run function with sequence
kable(t(format(out, scientific = F)), 
      col.names = c("Word Length",
                    "Number of Words",
                    "Without Repeats",
                    "Percent Same")) # drop scientific notation, transpose list, and output table
```

### Exercise 2.2.1

```{r}
15 * 14 * 13 * 12 #number of options for office positions, removing one for each election
# order of election shouldn't matter
```

### Exercise 2.2.2

```{r}
n <- c(1:5) #create sequence of "words". Here we want to see up to 5 letter words

f <- function(n,nw) { # modify function to include any "alphabet"
    now <- nw ^ n # calculate number of words
    wor <- factorial(nw) / (factorial(nw - n)) # calculate without repeats using permutations
    perc <- round(wor/now*100,0) # calculate percentage
    matrix(c(n, now,wor,perc))# output data
    
}
out <- sapply(n, f, nw = 24) #run function with greek alphabet and sequence
greek <- kable(t(format(out, scientific = F)))

outR <- sapply(n, f, nw = 33) #run function with russian alphabet and sequence
russian <- kable(t(format(outR, scientific = F)))
 
kable(list(t(out),t(outR)), # compare greek and russian alphabets
      col.names = c("Word Length",
                           "Number of Words",
                           "Without Repeats",
                           "Percent Same"),
      caption = "Greek (24 letters) vs. Russian Alphabet (33 letters)")
```

## 2.3 A Sequence of choices

### Exercise 2.3.1

```{r}
total <- factorial(36) / (factorial(36 - 6)) # use second equation from the previous function

cat("You have a 1 in", total, "chance to win!")
```
### Exercise 2.3.2
```{r}
n <- c(1:10) #create sequence of "words". Here we want to see up to 5 letter words

f <- function(n,nw) { # modify function to include any "alphabet"
    now <- nw ^ n # calculate number of words
    wor <- factorial(nw) / (factorial(nw - n)) # calculate without repeats using permutations
    perc <- round(wor/now*100,0) # calculate percentage
    matrix(c(n, now,wor,perc))# output data
    
}
out <- sapply(n, f, nw = 22)
kable(t(format(out, scientific = F)),
      col.names = c("Word Length",
                           "Number of Words",
                           "Without Repeats",
                           "Percent Same"),
      caption = "Hebrew Alphabet (22 letters)")
```


## 2.4 Factorials

### Exercise 2.4.1 - 2.4.4
```{r}
# The first exercise is taken care of for us :)
```


## 2.5 When order matters


